Maximising the Guaranteed Feasible Set for Uncertain MPC Schemes with Probabilistic Constraints

### 2016-10-26 draft.tex
Updated draft.
what happened

[ticket: X]

\providecommand{\conv}{\text{conv}}
\providecommand{\vol}{\text{vol}}
\providecommand{\spann}{\text{span}}
\providecommand{\C}{\mathcal C}
\providecommand{\E}{\mathcal E}
\providecommand{\W}{\mathcal W}
\providecommand{\V}{\mathcal V}
\providecommand{\X}{\mathcal X}
\providecommand{\Y}{\mathcal Y}
\providecommand{\U}{\mathcal U}
\providecommand{\R}{\mathcal R}
\providecommand{\T}{\mathcal T}
\providecommand{\D}{\mathscr D}
\providecommand{\PP}{\mathbb P}
\providecommand{\RR}{\mathbb R}
\providecommand{\bfa}[1]{\mathbf{#1}}

@@

\section{Introduction}
%
%
Determining positive invariant sets for uncertain linear systems has been discussed in the literature for decades, see e.g.~\cite{Kolmanovsky:1995,Kolmanovsky:1998,blanchini:2007}.
%
Properties of such sets and the algorithms to compute them have are known and documented in the literature.
%
In addition to pure analysis purposes, these sets are often used as terminal constraint sets in model predictive control formulations in order to guarantee the feasibility of the control scheme at all times, see e.g.~\cite{Mayne2014}.
%
To be able to apply the methods presented in~\cite{Kolmanovsky:1998} and~\cite{blanchini:2007} the modelled uncertainty has to be constrained to known convex set, in particular a polytope or a ellipsoid.
%
If constraints on the system do not have to be satisfied for all possible uncertainties but merely for a subset satisfying some probabilistic condition, sampling based methods can be applied, see e.g.~\cite{Margellos2014,Zhang2015}.
%
However, the sets determined based on sampled uncertainties are computationally expensive to obtain and guaranties can only be given in a stochastic sense, i.e. with a certain confidence level.
%
The method presented in~\cite{Zhang2015} is based on determining a certain set in which the probabilistic requirements are met and using that set for subsequent analysis.
%
This concept does not require sampling for simple probability distributions, instead we can explicitly parametrise an auxiliary set in which all probabilistic conditions are met.
%
Using such a parametrised set, we can optimise the size of the positive invariant set for the particular set in order to produce the largest positive invariant set while guaranteeing satisfaction of the probabilistic constraints, hence approximating the positive invariant set of the original problem.
%
Unlike sample based approaches this method inherently guarantees probabilistic constraint satisfaction.

@@

The following notation is used in the sequel: $A\oplus B = \{x:x=a+b,a\in A,b\in B\}$ denotes the Minkowski addition, $A\ominus B = \{x: x+b \in A\forall b\in B\}$ the Pontryagin difference, for a real set $C\subset\RR^d$ we denote the image of $C$ under the linear map $A:\RR^d\rightarrow\RR^d$ by $AC = \{y\in\RR^d:Ax=y,x\in C\}$ and the pre-image $A^{-1}C = \{y\in\RR^d:Ay\in C\}$.
%
We denote the probability of an event $E$ by $\PP\{E\}$, throughout this paper we assume the probability density to be uniform.

@@

The discrete time system we consider is given by
%
\begin{equation}
	x^+ = Ax + Bu + w
\end{equation}
%
with the state constraints $x\in\X = \{x\in\RR^d:E_ix\leq 1, i\leq M_\X\}$, the input~$u\in\U = \{u\in\RR^{q_\U}:F_i u\leq 1, i\leq M_\U\}$ and an uncertain variable $w\in\W = \{w\in\RR^{q_\W}:G_i w\leq 1,i\leq M_\W\}$.
%
We furthermore consider the auxiliary output variable 
%
\begin{equation}
	y = Cx + Du + v
\end{equation}
%
which is subject to a random variable $v\in\V = \{v\in\RR^{q_\Y}:\Gamma_i v\leq 1,i\in M_\V\}$ for which the probability distribution is assumed to be uniform.
%
The output is probabilistically constrained to satisfy
%
\begin{equation}\label{eq:probabilistic:constraint}
	\PP\{y\in\Y\}\geq p.
\end{equation}
%
with $\Y=\{y\in\RR^{q_\Y}:H_i y\leq h_i, i\leq M_\Y\}$.

Notice that for any $\V^\prime\subseteq\V$ with $\frac{\vol(\V^\prime)}{\vol(\V)}\geq p$ we have $\PP\{Cx+Du + v\in\Y\forall v\in\V^\prime\}\geq p$, where $Cx+Du+v\in\Y$ for all $v\in\V^\prime$ can easily be enforced by constraining $x$ and $u$.
%
Furthermore, by choosing $\V^\prime$ to be polytopic the equivalent constraints on $x$ and $u$ are polytopic and can easily be determined.
%
In the sequel of this paper we discuss how~$\V^\prime$ can be chosen in order to obtain the largest guaranteed n-step controllable set for a given target set, guaranteed positive invariant set.

@@

\begin{thm}
Iteration~\eqref{eq:MRPI:sequence} terminates after a finite number of iterations~$N$, such that $X_{N+1}=X_N$ and $X_N=X^\infty$.
\end{thm}
%
\begin{pf}
Do we need one?!
\end{pf}

@@

>>>>

\providecommand{\conv}{\text{conv}}
\providecommand{\vol}{\text{vol}}
\providecommand{\spann}{\text{span}}
\providecommand{\B}{\mathcal B}
\providecommand{\C}{\mathcal C}
\providecommand{\E}{\mathcal E}
\providecommand{\W}{\mathcal W}
\providecommand{\V}{\mathcal V}
\providecommand{\X}{\mathcal X}
\providecommand{\Y}{\mathcal Y}
\providecommand{\Z}{\mathcal Z}
\providecommand{\U}{\mathcal U}
\providecommand{\R}{\mathcal R}
\providecommand{\T}{\mathcal T}
\providecommand{\D}{\mathscr D}
\providecommand{\PP}{\mathbb P}
\providecommand{\RR}{\mathbb R}
\providecommand{\bfa}[1]{\mathbf{#1}}

@@

\section{Introduction}
%
%
Determining positive invariant sets for uncertain linear systems has been discussed in the literature for decades, see e.g.~\cite{Kolmanovsky:1995,Kolmanovsky:1998,blanchini:2007}.
%
Properties of such sets and the algorithms to compute them have are known and documented in the literature.
%
In addition to pure analysis purposes, these sets are often used as terminal constraint sets in model predictive control formulations in order to guarantee the feasibility of the control scheme at all times, see e.g.~\cite{Mayne2014}.
%
To be able to apply the methods presented in~\cite{Kolmanovsky:1998} and~\cite{blanchini:2007} the modelled uncertainty has to be constrained to known convex set, in particular a polytope or a ellipsoid.
%
If constraints on the system do not have to be satisfied for all possible uncertainties but merely for a subset then the problem becomes a chance constrained problem, see e.g.~\cite{Kall2010}.
%
Control formulations of this kind can be addressed using a range of techniques including scenario optimization, see e.g.~\cite{Calafiore2010}, and sampling based methods, e.g.~\cite{Margellos2014,Zhang2015}.
%
However, the sets determined based on sampled uncertainties are computationally expensive to obtain and guaranties can only be given in a stochastic sense, i.e. with a certain confidence level.
%
The method presented in~\cite{Zhang2015} is based on determining a certain set in which the probabilistic requirements are met with a given confidence and using that set for subsequent analysis.
%
This concept does not require sampling for simple probability distributions, instead we can explicitly parametrise an auxiliary set in which all probabilistic conditions are met.
%
Using such a parametrised set, we can optimise the size of the positive invariant set for the particular set in order to produce the largest positive invariant set while guaranteeing satisfaction of the probabilistic constraints, hence approximating the positive invariant set of the original problem.
%
Unlike sample based approaches this method inherently guarantees probabilistic constraint satisfaction.

A similar principle of using sets of a set probability measure is used in p-efficiency theory, see~\cite{Dentcheva2009}, where the distribution is not required to be simple but the structure of the chance constraints is restrictive.

@@

The following notation is used in the sequel: $A\oplus B = \{x:x=a+b,a\in A,b\in B\}$ denotes the Minkowski addition, $A\ominus B = \{x: x+b \in A\forall b\in B\}$ the Pontryagin difference, for a real set $C\subset\RR^d$ we denote the image of $C$ under the linear map $A:\RR^d\rightarrow\RR^d$ by $AC = \{y\in\RR^d:Ax=y,x\in C\}$ and the pre-image $A^{-1}C = \{y\in\RR^d:Ay\in C\}$.
%
We denote the probability of an event $E$ by $\PP\{E\}$, throughout this paper we assume the probability density to be uniform.
%
The set $\B_P(r) = \{x\in\RR^d:x^TPx\leq r^2\}$ denotes the $P$-ball centred at the origin, for a matrix $\Gamma\in\RR^{m\times n}$ the notation $\Gamma x\leq \bfa{1}$ is to be understood element-wise, where $\bfa{1}$ is a vector of ones in $\RR^m$.

@@

The discrete time system we consider is given by
%
\begin{equation}
	x^+ = Ax + Bu + w
\end{equation}
%
with the state constraints $x\in\X = \{x\in\RR^d:E_ix\leq 1, i\leq M_\X\}$, the input~$u\in\U = \{u\in\RR^{q_\U}:F_i u\leq 1, i\leq M_\U\}$ and an uncertain variable $w\in\W = \{w\in\RR^{q_\W}:G_i w\leq 1,i\leq M_\W\}$.
%
We furthermore consider the auxiliary output variable 
%
\begin{equation}
	y = Cx + Du + v
\end{equation}
%
which is subject to a random variable $v\in\V = \{v\in\RR^{q_\Y}:\Gamma_i v\leq 1,i\in M_\V\}$ for which the probability distribution is assumed to be uniform.
%
The output is probabilistically constrained to satisfy
%
\begin{equation}\label{eq:probabilistic:constraint}
	\PP\{ v \in \mathcal{V} : y = Cx + Du + v \in \mathcal{Y} \}\geq p.
\end{equation}
%
with $\Y=\{y\in\RR^{q_\Y}:H_i y\leq h_i, i\leq M_\Y\}$.

We not use the following proposition.
%
\begin{prop}
Let $v\in\V$ be a uniformly distributed random variable, let $\V^\prime\subseteq\V$ such that $\vol(\V^\prime)\geq p$, then every $q\in\Z\ominus\V^\prime$ satisfies $\PP\{q+v\in\Z\}\geq p$.
\end{prop}
%
\begin{pf}
%
Using that $\PP\{\{q\}\oplus\V^\prime\subseteq\Z\} = \PP\{\V^\prime\} = \vol(\V^\prime)$ for all $q\in\Z\ominus\V^\prime$ we have
%
\begin{equation}\begin{split}
&\PP\{q+v\in\Z\} = \PP\{\{q\}\oplus\V\subseteq\Z\} = \\
&\PP\{\{q\}\oplus(\V^\prime\cup(\V\setminus\V^\prime))\subseteq\Z\} = \\
&\PP\{\{q\}\oplus\V^\prime\subseteq\Z\} + \PP\{\{q\}\oplus(\V\setminus\V^\prime)\subseteq\Z\}\geq \\
&\PP\{\{q\}\oplus\V^\prime\subseteq\Z\} =\vol(\V^\prime)\geq p.
\end{split}\end{equation}
%
\end{pf}
%
Notice that for polytopic $\Z$ and $\V^\prime$ the set $\Z\ominus\V^\prime$ can be computed explicitly and is itself polytopic.

In the sequel of this paper we discuss how~$\V^\prime$ can be chosen in order to obtain the largest guaranteed n-step controllable set for a given target set, guaranteed positive invariant set.

@@

\begin{thm}
Let $\Gamma$ be a map such that $\Gamma x\leq\bfa{1}$ and $-\Gamma x\leq\bfa{1}$ for all $x\in\X$ and let $((A+BK),\Gamma)$ be observable, then iteration~\eqref{eq:MRPI:sequence} terminates after a finite number of iterations~$N$, such that $X_{N+1}=X_N$ and $X_N=X^\infty$.
\end{thm}
%
\begin{pf}
%
The observability of $((A+BK),\Gamma)$ implies that $X_d$ is compact, since $X_{d-1}\subseteq\bigcap_{k\leq d-1}(A+BK)^{-k}\X\subseteq\{x:\pm\Gamma x\leq\bfa{1}\wedge\dots\wedge\pm\Gamma(A+BK)^{d-1} x\leq\bfa{1}\} = \{x:\pm\Gamma Ox\leq \bfa{1}\}$ for the observability matrix~$O$ which has full rank.

Let~$P$ be such that $(A+BK)^TP(A+BK)\leq\rho^2P$ holds for $\rho$ denoting the spectral radius of $A+BK$.
%
To see that $\E_k$ grows exponentially let $\W\subseteq\B_P(r)$ then
%
\begin{equation}\begin{split}
	&\bigoplus_{i=0}^{k-1}(A+BK)^i\W \subseteq \bigoplus_{i=0}^{k-1}(A+BK)^i\B_P(r)\\
	&\subseteq\bigoplus_{i=0}^{k-1}\B_P(\rho^i r) = \B_P(\sum_{i=0}^{k-1}\rho^i r)\subseteq\B_P(\frac{r}{1-\rho})
\end{split}\end{equation}
%
i.e. $\E_k$ consists of an expanding minuend and a bounded subtrahend.
%
Therefore $\E_k$ expands exponentially, since $X_{d-1}$ is bounded there exists an integer~$N$ such that $X_{N}\cap\E_{N+1}=X_N$ hence $X_N=X^\infty$.
%
\end{pf}

@@

\begin{rem}
The goal of reformulating a chance constrained model predictive control setup using the proposed method would be to minimise the introduced conservatism.
%
To achieve the best possible performance while guaranteeing constraint satisfaction a closed loop min-max model predictive control problem would be formulated, see e.g.~\cite{Lee:1997}.
%
This means that at each stage the decision variables are constrained to be such that the successor state is guaranteed to lie on a feasible trajectory to the target set.
%
So that if a chance constrained model predictive control problem is reformulated as a deterministically constrained problem guaranteeing probabilistic feasibility, the sets $C_n(\T)$ would play the role of stage constraints with the target set $\T=X^\infty$ to guarantee recursive feasibility.
\end{rem}
